\documentclass[11pt,letterpaper]{article}
\usepackage{preamble}

 
\title{Homework Assignment 04 \\
    \small Due: September 27, 2019}

\begin{document}
\maketitle



\section{(inverse) homomorphism}
\label{sec:1}
By the last question of the previous homework, 
$h^{-1}(L) = (a+ba)^*$ 
[Problem Statement: \textit{Let $L=(0+10)^*$, $h(a)=00$, and $h(b)=01$. What is $h^{-1}(L)$?}]. 
Passing $h^{-1}(L)$ through $h$ is trivial as $h(a)$ and $h(b)$ are known. Hence, $h(h^{-1}(L)) = (h(a) + h(b)h(a))^* = (00+0100)^*$. \\
$L$ is described as $(0+10)^*$. Hence, the first part of the answer above ($(00)^*$) is contained within $L$. Moreover, the second part, $(0100)^*$ is contained within $L$ as well. \small{If $m=0$ and $n=10$, $L=(m+n)^*$. $00$ is the same as $(mm)^*$ and $(0100)^*$ is the same as $(mnm)^*$.} 
Hence, the result above, $h(h^{-1}(L)) = (00+0100)^*$ is contained within $L = (0+10)^*$.

\section{which is larger?}
\begin{center}
    $h(h^{-1}(L)) \subseteq	L$
\end{center}
\paragraph{cheating}
The wording of the last question of the previous homework ``\textit{Show that your answer for $h(h^{-1}(L))$ is contained in $L$}" suggests the above relation.
\paragraph{informal}
From \hyperref[sec:1]{Section 1}, we see that $h(h^{-1}(L)) = (00+0100)^*$ which is contained in $L=(0+10)^*$. However, $L$ contains many more strings which are not in $h(h^{-1}(L))$ such as, $0, 10, 010, 0010, \cdots$ Based on this \textit{evidence}, $|L| > |h(h^{-1}(L))|$ and hence the above relation holds true.

\section{where are the proofs?!}
Let the theorem be denoted by $a$ and the proof be denoted by $\^{a}$. Then we can create a homomorphism such that all possible proofs are inserted randomly between the theorems. More formally, $h_1(a) = a \wedge h_1(\^{a}) = a$. Then $h^{-1}(book)$ gives the set of all strings where $\^{a}$ appears at random intervals, covering all possibilities (ie, a set containing $2^n$ strings). \\
Now, the required regular set is a set containing all the theorems with the proofs. Intersecting the book (passed through $h_1$) with the language of this regular set yields in all strings where the theorem is followed by the proof. Formally, $(h_1^{-1}(book)) \cap (a\^{a})^*$ is the resulting set. \\
Finally, let us create another homomorphism which maps all the theorems to themselves and discards all the proofs. We can describe such a transformation by $h_2(a) = a \wedge h_2(\^{a}) = \epsilon$. Passing the set obtained above through this transformation, results in the set of all theorems but no proofs. \\
Therefore we can extract the set of all theorems ($a$) by the following operation,
\begin{center}
    $h_2(h_1^{-1}(book) \cap (a\^{a})^*)$ where, \\
    $h_1(a) = a \wedge h_1(\^{a}) = a$, \\
    $h_2(a) = a \wedge h_2(\^{a}) = \epsilon$
\end{center}

\section{closure has come to me, myself}
\begin{center}
    $M = (Q,\Sigma,\delta,q_0,F)$
\end{center}
We can redirect the language to be passed under a described homomorphism $h$ and then that transformed set of strings to go through the machine $M$. We can finally also check if the machine accepts the initial string $x$ as well. On doing so, we construct a machine as shown in  \hyperref[fig:4a]{Figure 1}. \\
We can remove the second machine and redirect the output of $h$ in \hyperref[fig:4a]{Figure 1}. This allows us to remove the $and$ state as well. This construction looks like \hyperref[fig:4b]{Figure 2}.
Therefore an automaton that accepts $h(L(M))$ is the same as that which accepts $L(M)$. Hence, $M = (Q,\Sigma,\delta,q_0,F)$ describes a machine that accepts $h(L(M))$. \pagebreak
\begin{figure}[!htp]
\label{fig:4a}
\centering
\begin{tikzpicture}
% States %
\node[state, initial] at (-4,0.5) (S) {};
\node[state] at (-1,1) (h) {$h$};
\node[] at (0,0) (M) {$M$};
\node[] at (1,1) (Mh) {$M$};
\node[state] at (3,0.5) (A) {$and$};
\node[state, accepting] at (6,0.5) (F) {};
% Edges %
\draw [->]
    (S) edge[above] node{x} (h)
    (S) edge[above] node{x} (M)
    (h) edge[above] node{h(L)} (Mh)
    (M) edge[above] node{1} (A)
    (Mh) edge[above] node{1} (A)
    (A) edge[above] node{$1\wedge1$} (F);
\end{tikzpicture}
\caption{Closure under homormorphism}
\end{figure}
\begin{figure}[!htp]
\label{fig:4b}
\centering
\begin{tikzpicture}
% States %
\node[state, initial] at (-3,0) (S) {};
\node[state] at (0,2) (h) {$h$};
\node[] at (0,0) (M) {$M$};
\node[state, accepting] at (3,0) (F) {};
% Edges %
\draw [->]
    (S) edge[above] node{x} (h)
    (S) edge[above] node{x} (M)
    (h) edge[right] node{h(L)} (M)
    (M) edge[above] node{1} (F);
\end{tikzpicture}
\caption{Closure under homormorphism (simple)}
\end{figure}

\section{closure under inverse substitution}
The given regular set, $R$, describes a language consisting of the symbols in $\Sigma$. Let $s:\Sigma \rightarrow \Delta^*$ such that $s(\epsilon) = \epsilon$ and $s(xa) = s(x)s(a);$ $\forall x \in \Sigma^*, a \in \Sigma$. We now define a machine $M$ that accepts the regular set $R$ as 
\begin{center}
    $M_1=(Q,\Sigma,\delta_1,q_0,F)$ 
\end{center}
To prove that $s^{-1}(R)$ is a regular set, we construct another machine that accepts $s^{-1}(R)$ as $M_2=(Q,\Delta,\delta_2,q_0,F)$, where,
\begin{center}
    $\delta_2(q,x)=\delta_1(q,s(x))$.
\end{center}
We can prove the above relation using induction on $|x|$. 
\paragraph{Hypothesis} $\delta_2(q,x)=\delta_1(q,s(x))$
\paragraph{Basis} When $x=\epsilon$, $s(x)=\epsilon$ and hence, $\delta_2(q_0,\epsion) = \delta_1(q_0,s(\epsilon)) = \delta_1(q_0,\epsilon) = q_0$
\paragraph{Induction} Let $x=wa$. Then, $\delta_2(q_0,x) = \delta_2(\delta_2(q_0,w),a)$; By the hypothesis, $\delta_2(\delta_2(q_0,w),a)$ $=$ $\delta_1(\delta_1(q_0,s(w)),s(a))$ $=$ $\delta_1(q_0,s(w)s(a))$ $=$ $\delta_1(q_0,s(wa))$ $=$ $\delta_1(q_0,s(x))$
\end{document}