\documentclass[11pt,letterpaper]{article}
\usepackage{preamble}

\title{Homework Assignment 05 \\
    \small Due: October 4, 2019}

\begin{document}
\maketitle

\section{same same but different}
Regular expressions are one way to represent the strings in a language. The same set of strings can be represented by an automaton as well. Hence, the easiest way to check if two expressions are equivalent, we should construct a automaton for each expression and then minimize the constructed automaton. As there is a unique minimized automaton for every set, if the obtained minimized automata are the same, then the regular expressions that we started from denote the same set.

\section{homomorphism}
Let the given set be denoted by $S$. Therefore, $S = {wcw | w \in (a+b)^*}$. Consider a homomorphism $h_1$ such that $h_1(a+b+c)=(a+b+c)$ and $h_1(\^{a}+\^{b}+\^{c})=(a+b+c)$ (obviously, the letters match on both sides of the equation). Therefore, the inverse of transformation $h_1$ (ie, $h_1^{-1}(S)$ would yield the set that contains $2^n$ strings with the \textit{hats} randomly inserted in the strings in $S$. \\
If that set is intersected with the set described by, $(a+b+c)^n(\^{a}+\^{b}+\^{c})^n$, we discard all the strings where the number of letters is not equal to number of letter hats. Finally, a second homomorphism $h_2$ that  transforms $\Sigma = {a,b,c}$ to $\Delta = {0,1}$ in the following ways, $h_2(a+b+c)=0$ and $h_2(\^{a}+\^{b}+\^{c})=1$, transforms the set obtained via intersection to the set $\{0^n1^n\}$. Hence, 
\begin{center}
    $h_2(h_1^{-1}(S) \cap (a+b+c)^n(\^{a}+\^{b}+\^{c})^n) = \{0^n1^n\}$
\end{center}

\section{pumping lemma}
The given set describes a language defined as $L=(a+b)^*c(a+b)^*$. Let us consider a string $v \in L$ such that $v = xyz$ and $|xy| \leq |v|$ where $x=(a+b)^*,y=c,z=(a+b)^*$. \\
The pumping lemma states that if $v=xyz \in L$ and $L$ is regular, then $\exists i \geq 0$ such that $xy^iz \in L$. By definition, $L$ can contain only one $c$ and hence, the claim will not hold true whenever $i \neq 1$. As $xy^iz \notin L, i > 1$, $L$ is not regular. Hence, the given set is not regular.

\section{are you not compelled?}
Similar to the previous question, let us split the strings in our set (read: language) into the form $xyz$ where $x=(01)^n0^*, y=010, z=(10)^n1^*(\epsilon+1)$. In other words, let $y$ be the troublesome $010$ and $x$ and $z$ contain an equal number of $01$s and $10$s respectively. The string may contain contain other $0$s and $1$s and that is taken care of with the $0^*$ in $x$ and the $1^*(\epsilon+1)$ in $z$. The variation in length is also taken care of with the $a^*, a={0,1}$ and $(\epsilon+1)$. Hence, the given set appears to be regular under preliminary thinking. \\
Note that the above expressions do not consider all possibility of strings. However, cases can be taken and it can be showed similarly that $L$ is indeed regular.
\pagebreak
\section{minimization}
The given automata can be drawn as \hyperref[fig:5a]{below}. If you notice, states $A$ and $D$ and $B$ and $E$ can be merged to minimize the automata. The \hyperref[fig:5a]{minimized automata} contains only three states, namely - $A$, $B$, and $C$. The bidirectional flow between the states ($A$-$D$, $B$-$E$) get converted to loops. The automaton accepts all strings that contain at least one $0$ and ends with a $1$.
\begin{figure}[!ht]
\label{fig:5a}
\centering
\begin{tikzpicture}
% States %
\node[state, initial] at (-2,0) (A) {A};
\node[state] at (0,0) (B) {B};
\node[state, accepting] at (2,0) (C) {C};
\node[state] at (-2,2) (D) {D};
\node[state] at (0,2) (E) {E};
% Edges %
\draw [->]
    (A) edge[above] node{0} (B)
    (A) edge[left] node{1} (D)
    (B) edge[right] node{0} (E)
    (B) edge[bend left, above] node{1} (C)
    (C) edge[bend left, below] node{0} (B)
    (C) edge[loop right] node{1} (C)
    (D) edge[above] node{0} (E)
    (D) edge[] node{} (A)
    (E) edge[] node{} (B)
    (E) edge[bend left, above] node{1} (C);
\end{tikzpicture}
\caption{Given automata}
\end{figure}
\begin{figure}[!ht]
\label{fig:5b}
\centering
\begin{tikzpicture}
% States %
\node[state, initial] at (-2,0) (A) {A};
\node[state] at (0,0) (B) {B};
\node[state, accepting] at (2,0) (C) {C};
% Edges %
\draw [->]
    (A) edge[above] node{0} (B)
    (A) edge[loop above] node{1} (A)
    (B) edge[loop above] node{0} (B)
    (B) edge[bend left, above] node{1} (C)
    (C) edge[bend left, below] node{0} (B)
    (C) edge[loop above] node{1} (C);
\end{tikzpicture}
\caption{Minimized automata}
\end{figure}


\end{document}